---
title: "TSGroup3Latency"
author: "SLP"
date: "Thursday, November 13, 2014"
output: html_document
---

This script takes as its input reduced OR raw data from ABETII, exported as csv files, with one file corresponding to the performance of a single mouse during a single session.  Reduced data must contain at least the events "Display Image", "Initiate ChoicePhase", "Display Choice Images", "Correct", and "Incorrect".  Presence of additional events in raw files is not problematic and will be skipped over if not matching any needed events.  However, if any information is present in the csv file prior to the header, this must be removed manually prior to analysis.

Files to use are pulled from a subfolder "StimulusConfigData" under the directory: 

`r getwd() ` 

Complete sets of mice are not required, as each session is processed individually.

Warning: Only raw files from MatchLocation and NonMatchLocation schedules can be processed.  Pretraining (Must Touch, PunishIncorrect) or object-based schedules will crash.


```{r}
filestouse <- {}
for(mouse in 1:12){
    filepattern <- paste("Mouse", mouse, "raw.csv", sep = "")
    filestouse <- append(filestouse, list.files(path = "../StimulusConfigData/", pattern = filepattern))
}
```

Data from the following dates will be analyzed:
```{r}
datestouse <- unique(as.Date(gsub("Mouse[1-12]raw.csv", "", filestouse), format = "%Y%m%d"))
print(datestouse)
```

Read in schedule info from data summary sheet, same as used in daily reporting script.  This sheet MUST be current up to the last date present in the raw csv files.  
```{r}
# Read in manually entered data for date, mouseid, and schedule
schedinfo <- read.csv(file = "../Data Summary.csv", header = TRUE, sep = ",", stringsAsFactors = FALSE)
#Reformat scheduleinfo dates
schedinfo$Date <- as.Date(schedinfo$Date, format = "%m/%d/%Y")
schedtrim <- schedinfo[schedinfo$Date %in% datestouse ,1:4]
#Mice are not integers
schedtrim$Mouse <- as.character(schedtrim$Mouse)
#Print extracted info for first day
print(head(schedtrim,12))
#Print extracted info for last day
print(tail(schedtrim,12))
```

Generate empty data frame for storing reduced data
```{r}
dfkeep <- data.frame(date = {}, mouse = {}, schedule = {}, sampleinit = {}, sampleresponse = {}, choiceinit = {}, choiceresponse = {}, correct = {})
```

### Loop through individual files and bind data across mice and days.

```{r}
for(fil in filestouse){
    ###Read in raw data for a single session, extract mouse number, session date, and schedule
    tempfile <- read.csv(file = paste("../StimulusConfigData/",fil,sep=""), header = TRUE, stringsAsFactors = FALSE)
    numrows <- length(tempfile$Item_Name)
    mouseid <- as.integer(gsub("^[0-9]*(Mouse)|raw.csv","",fil))
    sessiondate <- as.Date(gsub("Mouse[1-12]raw.csv","", fil), format="%Y%m%d")
    sessionsched <- schedtrim[(schedtrim$Date == sessiondate) & (schedtrim$Mouse == mouseid),4]    

    
    ###generate empty data frame with number of rows based on number of trials initiated
    dfrows <- sum(tempfile$Item_Name == "Display Image")
    latdf <- data.frame(date = rep(sessiondate, dfrows), mouse = mouseid, schedule = sessionsched, sampleinit = "", sampleresponse = "", choiceinit = "", choiceresponse = "", correct = "")
    
    ###compare number of initiated to completed trials. if trial incomplete then set all remaining timestamps for final trial to 3600
    initiated1 <- sum(tempfile$Item_Name == "Display Image")
    responded1 <- sum(tempfile$Item_Name == "Initiate ChoicePhase")
    initiated2 <- sum(tempfile$Item_Name == "Display Choice Images")
    if(initiated1 != responded1){padresponse1<- TRUE} else {padresponse1<-FALSE}
    if(initiated1 != initiated2){padinitiated2 <- TRUE} else {padinitiated2<-FALSE}
    
    ###populate data frame with data from current file
    latdf$sampleinit <- tempfile[tempfile$Item_Name == "Display Image",1]
    if(padresponse1 == TRUE){
        latdf$sampleresponse <- append(tempfile[tempfile$Item_Name == "Initiate ChoicePhase",1],3600)
    } else {
        latdf$sampleresponse <- tempfile[tempfile$Item_Name == "Initiate ChoicePhase",1]
    }
    if(padinitiated2 == TRUE){
        latdf$choiceinit <- append(tempfile[tempfile$Item_Name == "Display Choice Images",1], 3600)
    } else {
        latdf$choiceinit <- tempfile[tempfile$Item_Name == "Display Choice Images",1]
    }
    corlat <- tempfile[tempfile$Item_Name == "Correct",1]
    incorlat <- tempfile[tempfile$Item_Name == "Incorrect",1]
    choicelat <- sort(append(corlat, incorlat))
    if(length(choicelat) != initiated1) {choicelat <- append(choicelat, 3600)}
    latdf$choiceresponse <- choicelat
    latdf$correct <- (latdf$choiceresponse %in% corlat)

    ###bind data from individual file to all previous files
    dfkeep <- rbind(dfkeep, latdf)
}
```
    
###Determine interresponse latencies
```{r}
    dfkeep$sampleresponselat <- dfkeep$sampleresponse - dfkeep$sampleinit
    dfkeep$choiceinitlat <- dfkeep$choiceinit - dfkeep$sampleresponse
    dfkeep$choiceresponselat <- dfkeep$choiceresponse - dfkeep$choiceinit
library(plyr)
###summarize response latencies, split by mouse and schedule
    lattest<- ddply(dfkeep, .(mouse, schedule), summarize, sampleresponsemean = mean(sampleresponselat), sampleresponsemedian = median(sampleresponselat), choiceinitmean = mean(choiceinitlat), choiceinitmedian = median(choiceinitlat), choiceresponsemean = mean(choiceresponselat), choiceresponsemedian = median(choiceresponselat), totaltrials = length(sampleinit))
```

###Single schedules
```{r fig.height = 8, fig.width = 8}
latsingle <- lattest[grep("Single", lattest$schedule),]
library(ggplot2)
###graph 3 main response latency measures over single schedules
g <- ggplot(latsingle) + geom_point(aes(x = "1:SampleResponse", y = sampleresponsemedian, color = schedule), position = position_jitter(w = 0.1), size = 4) + geom_point(aes(x = "2:ChoiceInit", y = choiceinitmedian, color = schedule), position = position_jitter(w = 0.1), size = 4) + geom_point(aes(x = "3:Choice Response", y = choiceresponsemedian, color = schedule), position = position_jitter(w = 0.1), size = 4) + xlab(label = "") + ylab("Median Latency (s)")
print(g)
latdouble <- lattest[grep("Double", lattest$schedule),]
###graph 3 main response latency measures over double schedules
g <- ggplot(latdouble) + geom_point(aes(x = "1:SampleResponse", y = sampleresponsemedian, color = schedule), position = position_jitter(w = 0.1), size = 4) + geom_point(aes(x = "2:ChoiceInit", y = choiceinitmedian, color = schedule), position = position_jitter(w = 0.1), size = 4) + geom_point(aes(x = "3:Choice Response", y = choiceresponsemedian, color = schedule), position = position_jitter(w = 0.1), size = 4) + xlab(label = "") + ylab("Median Latency (s)")
print(g)
```