---
title: "TSGroup3Latency"
author: "SLP"
date: "Thursday, November 13, 2014"
output: html_document
---

This script takes as its input reduced OR raw data from ABETII, exported as csv files, with one file corresponding to the performance of a single mouse during a single session.  Reduced data must contain [?].  Presence of additional events in raw files is not problematic and will be skipped over if not matching any needed events.  However, if any information is present in the csv file prior to the header, this must be removed manually prior to analysis.

Files to use are pulled from a subfolder "raw" under the directory: 

`r getwd() ` 

Complete sets of mice are not required, as each session is processed individually.


```{r}
filestouse <-"20141112Mouse3raw.csv"
#for(mouse in 1:12){
#    filepattern <- paste("Mouse", mouse, "raw.csv", sep = "")
#    filestouse <- append(filestouse, list.files(path = "../StimulusConfigData/", pattern = filepattern))
#}
```

Data from the following dates will be analyzed:
```{r}
datestouse <- unique(as.Date(gsub("Mouse[1-12]raw.csv", "", filestouse), format = "%Y%m%d"))
print(datestouse)
```

Read in schedule info from data summary sheet, same as used in daily reporting script.
```{r}
# Read in manually entered data for date, mouseid, and schedule
schedinfo <- read.csv(file = "../Data Summary.csv", header = TRUE, sep = ",", stringsAsFactors = FALSE)
#Reformat scheduleinfo dates
schedinfo$Date <- as.Date(schedinfo$Date, format = "%m/%d/%Y")
schedtrim <- schedinfo[schedinfo$Date %in% datestouse ,1:4]
#Mice are not integers
schedtrim$Mouse <- as.character(schedtrim$Mouse)
#Print extracted info for first day
print(head(schedtrim,12))
#Print extracted info for last day
print(tail(schedtrim,12))

#Generate list of schedules contained in data range of files
locationsched <- c("MatchLocationSingle", "NonMatchLocationSingle", "MatchLocationDouble", "NonMatchLocationDouble")
pretrainingsched <- c("Must Touch", "PunishInc")
```

Generate empty data frame for storing reduced data
{r}
dfkeep <- data.frame(date = {}, mouse = {}, schedule = {}, correctgrid = {}, samplegrid = {}, correct = {}, incorrect = {}, totaltrials = {})
totalskeep <- data.frame(date = {}, mouse = {}, schedule = {}, correctgrid = {}, samplegrid = {}, correcttotal = {}, incorrecttotal = {}, trialcount = {}, percent = {})

```

### Loop through individual files and extract mouse number, session date, and events from each file

```{r}
###Keep track of whether current file marks a switch to new mouse or schedule
currentmouse <- 0
currentsched <- ""

for(fil in filestouse){
        
    tempfile <- read.csv(file = paste("../StimulusConfigData/",fil,sep=""), header = TRUE, stringsAsFactors = FALSE)
    numrows <- length(tempfile$Item_Name)
    mouseid <- as.integer(gsub("^[0-9]*(Mouse)|raw.csv","",fil))
    if(mouseid != currentmouse) {
        totaltrials <- 0
        currentmouse <- mouseid
    }
    sessiondate <- as.Date(gsub("Mouse[1-12]raw.csv","", fil), format="%Y%m%d")
    sessionsched <- schedtrim[(schedtrim$Date == sessiondate) & (schedtrim$Mouse == mouseid),4]    
    if(sessionsched != currentsched) {
        totaltrials <- 0
        currentsched <- sessionsched
        }
    
    ###generate empty data frame with number of rows based on number of correct grid positions set in file
    dfrows <- sum(tempfile$Item_Name == "Display Image")
    latdf <- data.frame(date = rep(sessiondate, dfrows), mouse = mouseid, schedule = sessionsched, sampleinit = "", sampleresponse = "", choiceinit = "", choiceresponse = "")
    ###populate data frame with data from current file
    latdf$sampleinit <- tempfile[tempfile$Item_Name == "Display Image",1]
    latdf$sampleresponse <- tempfile[tempfile$Item_Name == "Initiate ChoicePhase",1]
    latdf$choiceinit <- tempfile[tempfile$Item_Name == "Display Choice Images",1]
    corlat <- tempfile[tempfile$Item_Name == "Correct",1]
    incorlat <- tempfile[tempfile$Item_Name == "Incorrect",1]
    latdf$choiceresponse <- sort(append(corlat, incorlat))
    latdf$correct <- ""
    latdf$correct <- (latdf$choiceresponse %in% corlat)
}

###START HERE 11/14
    ###bind data from individual file to all previous files
    dfkeep <- rbind(dfkeep, trackerdf)
    totaltrials <- trackerdf$totaltrials[dfrows]
    library(plyr)
    ###summarize correct and incorrect responses session and bind to all previous sessions
    dftotals <- ddply(trackerdf, .(date, mouse, schedule, correctgrid, samplegrid), summarize, correcttotal = sum(correct), incorrecttotal = sum(incorrect))
    dftotals$trialcount <- (dftotals$correcttotal + dftotals$incorrecttotal)
    dftotals$percent <- (dftotals$correcttotal/(dftotals$correcttotal + dftotals$incorrecttotal))
    totalskeep <- rbind(totalskeep, dftotals)
}
```
