---
title: "TSGroup3StimulusConfiguration"
author: "SLP"
date: "Tuesday, November 04, 2014"
output: html_document
---

This script takes as its input reduced data from ABETII, exported as csv files, with one file corresponding to 
the performance of a single mouse during a single session.  Reduced data must be queried for "Correct_Grid_Position", "SampleImage", "Correct", and "Incorrect".  If any information is present in the csv file prior to the header, this must be removed manually prior to analysis.

Files to use are pulled from a subfolder "raw" under the directory: 

`r getwd() ` 

and all schedule information (NonMatch vs. NonMatchEasy-Hard vs. Match) is currently ignored.  Complete sets of mice are not required, as each session is processed individually.

```{r}
filestouse <- list.files(path = "../StimulusConfigData/", pattern = "raw.csv")
datestouse <- unique(as.Date(gsub("Mouse[1-12]raw.csv", "", filestouse), format = "%Y%m%d"))
dfkeep <- data.frame(date = {}, mouse = {}, grid = {}, image = {}, correct = {}, incorrect = {})
totalskeep <- data.frame(date = {}, mouse = {}, grid = {}, image = {}, correcttotal = {}, incorrecttotal = {}, trialcount = {}, percent = {})
```

Data from the following dates will be analyzed:
```{r}
print(datestouse)
```

### Loop through individual files and extract mouse number, session date, and events from each file

```{r}
for(fil in filestouse){
        
    ms1 <- read.csv(file = paste("../StimulusConfigData/",fil,sep=""), header = TRUE, stringsAsFactors = FALSE)
    numrows <- length(ms1$Item_Name)
    mouseid <- as.integer(gsub("^[0-9]*(Mouse)|raw.csv","",fil))
    sessiondate <- as.Date(gsub("Mouse[1-12]raw.csv","", fil), format="%Y%m%d")
        
    gridid <- 0
    imageid <- 0

    ###generate empty data frame with number of rows based on number of correct grid positions set in file
    dfrows <- (sum(ms1$Item_Name == "Correct_Grid_Position") - 1)
    trackerdf <- data.frame(date = rep(sessiondate,dfrows), mouse = rep(mouseid,dfrows), grid = rep(0,dfrows), image = 0, correct = 0, incorrect = 0, row.names = (1:dfrows))
    ### skip initial 2 rows where grid position and image are set to zero
    rowindex <- 1
    ###loop through each row of csv file and fill values in empty data frame according to event type
    for(row in 3:numrows) {
        if(ms1[row, 4] == "Correct_Grid_Position") {
            gridid <- ms1[row, 9]
            trackerdf$grid[rowindex] <- gridid
        }
        else if(ms1[row, 4] == "SampleImage") {
            imageid <-ms1[row, 9]
            trackerdf$image[rowindex] <- imageid
        }    
        else if(ms1[row, 4] == "Correct") {
            trackerdf$correct[rowindex] <- 1
            rowindex <- rowindex + 1
        }
        else if(ms1[row, 4] == "Incorrect") {
            trackerdf$incorrect[rowindex] <- trackerdf$incorrect[rowindex] + 1
        }
    }
    ###bind data from individual file to all previous files
    dfkeep <- rbind(dfkeep, trackerdf)
    library(plyr)
    ###summarize correct and incorrect responses session and bind to all previous sessions
    dftotals <- ddply(trackerdf, .(date, mouse, grid, image), summarize, correcttotal = sum(correct), incorrecttotal = sum(incorrect))
    dftotals$trialcount <- (dftotals$correcttotal + dftotals$incorrecttotal)
    dftotals$percent <- (dftotals$correcttotal/(dftotals$correcttotal + dftotals$incorrecttotal))
    totalskeep <- rbind(totalskeep, dftotals)
}
###Save output files to output directory
write.csv(dfkeep, file = paste("../output/", Sys.Date(),"rawcombooutput.csv",sep=""), row.names = FALSE)
write.csv(totalskeep, file = paste("..output/", Sys.Date(),".combooutput.csv", sep=""), row.names = FALSE)
```

###Is the proportion of correct and incorrect trials equal across all side and image combinations?

Equality of performance across stimulus combinations will be checked via Pearson's chi-squared test. Subsequently, each stimulus configuration will be checked separately to determine if responding differs from chance performance, assumed to be 50%.


```{r}
###Perform chi-squared tests for each image/side combo
###Note that initial chi-squared across groups compares independence of distributions, whereas
###chi-squared tests of individual rows compare correct vs. incorrect responding to a hypothetical 50% chance 
combotabs <- ddply(totalskeep, .(grid, image), summarize, totaltrials = sum(trialcount), correct = sum(correcttotal), incorrect = sum(incorrecttotal), meanalldays = (correct/(correct+incorrect)))
print(combotabs)
###Across all possible configurations,
print(chisq.test(combotabs[,4:5]))
```

**A large disparity exists in responding to different configurations.  A further breakdown appears below.**

```{r}
###Check individual pairings to determine if each combination is above or below chance
###Side 1 Image 1:
print(chisq.test(combotabs[1,4:5]))
###Side 1 Image 2:
print(chisq.test(combotabs[2,4:5]))
###Side 3 Image 1:
print(chisq.test(combotabs[3,4:5]))
###Side 3 Image 2:
print(chisq.test(combotabs[4,4:5]))
```

**All configurations except Grid 3 Image 2 (Right, Sample image O) differ significantly from chance.**

```{r}
###Ignore image, check grid
combotabs <- ddply(totalskeep, .(grid), summarize, totaltrials = sum(trialcount), correct = sum(correcttotal), incorrect = sum(incorrecttotal), meanalldays = (correct/(correct+incorrect)))
print(combotabs)
print(chisq.test(combotabs[,3:4]))

###Grid side 1
print(chisq.test(combotabs[1,3:4]))
###Grid Side 3
print(chisq.test(combotabs[2,3:4]))
```

**Mice perform better when the Correct_Grid_Position is equal to 1 (on the left).**

```{r}
###Ignore grid, check image
combotabs <- ddply(totalskeep, .(image), summarize, totaltrials = sum(trialcount), correct = sum(correcttotal), incorrect = sum(incorrecttotal), meanalldays = (correct/(correct+incorrect)))
print(combotabs)
print(chisq.test(combotabs[,3:4]))

###Image 1
print(chisq.test(combotabs[1,3:4]))
###Image 2
print(chisq.test(combotabs[2,3:4]))
```

**Mice perform slightly better when the sample image is 2 (O) compared to image 1 (X).**

```{r fig.height = 8}
###Output graph of percent correct across days for all stimulus combinations split by mouse
bymouse <- ddply(totalskeep, .(mouse, grid, image), summarize, totaltrials = sum(trialcount), correct = sum(correcttotal), incorrect = sum(incorrecttotal), meanalldays = (correct/(correct+incorrect)))
bymouse$trialcode <- rep(c("X-Left","O-Left","X-Right","O-Right"))
bymouse$trialcode <- factor(bymouse$trialcode, levels = c("X-Left","O-Left","X-Right","O-Right"))
library(ggplot2)
g <- ggplot(bymouse, aes(x = trialcode, y = meanalldays)) + geom_point(size = 4) + coord_cartesian(ylim = c(0,1)) + facet_wrap(~ mouse, ncol = 3)
print(g)
```

**Across mice, the side preference appears more robust and consistent than the image preference.**