---
title: "TSGroup3Report"
author: "SLP"
date: "Thursday, October 23, 2014"
output: html_document
---

Touchscreen Group 3 Performance
========================================================

```{r set-options, echo = FALSE, cache = FALSE}
options(width = 100)
```

Analysis requires data from 2 sources, and formatting is restricted.

1. ABETII event total output: Event totals from each daily session must be present in a folder called "raw" that resides in the parent directory of this file.  Session files must be named as |4DigitYear||2DigitMonth||2DigitDay||Set1-3|_Group3.csv.  Verify manually that session info is deleted from the header of the csv files.  Headers should appear in row 2.

2. Schedule sheet: A csv file Data Summary.csv containing the Date, RunDay, Mouse, and Schedule info for each session must be in the working directory.  These must be updated manually to reflect decisions to move mice between schedules.  For every existing event total in 1, corresponding data MUST exist for that date in the Data Summary sheet.  If there are any days missing on the Data Summary sheet, it will crash.

Mouse identities in the ABETII event total sheet are assumed to be, from leftmost to rightmost column:

- Set 1: 1 to 4
- Set 2: 5 to 8
- Set 3: 9 to 12

Any changes to these identities (e.g. due to box failures that require running elsewhere) should be manually adjusted in the output files themselves and noted prior to analysis.

Prior to Punish Incorrect, performance will appear as 100% because there's no real way to get these trials wrong.  For all later schedules, percentall indicates the total percentage of correct trials, including correction trials.

The following R packages are required:

- base
- plyr
- ggplot2
- RColorBrewer

### Take inventory of files in raw directory and extract dates to report on.

```{r}
# determine files to use
filestouse <- list.files(path = "../raw/", pattern = "Group3.csv")
# Strip dates and set information from raw output files
datestouse <- unique(as.Date(gsub("Set[1-3]_Group3.csv|Set[1-3].csv", "", filestouse), format = "%Y%m%d"))
print(filestouse)
print(datestouse)
```

### Extract schedule information from manually updated Data Summary sheet.

```{r}
# Read in manually entered data for date, mouseid, and schedule
schedinfo <- read.csv(file = "../Data Summary.csv", header = TRUE, sep = ",", stringsAsFactors = FALSE)
#Reformat scheduleinfo dates
schedinfo$Date <- as.Date(schedinfo$Date, format = "%m/%d/%Y")
schedtrim <- schedinfo[schedinfo$Date %in% datestouse ,1:4]
#Mice are not integers
schedtrim$Mouse <- factor(as.character(schedtrim$Mouse), levels = c("1","2","3","4","5","6","7","8","9","10","11","12"))
#Print extracted info for first day
print(head(schedtrim,12))
#Print extracted info for last day
print(tail(schedtrim,12))

#Generate list of schedules contained in data range of files
schedulestouse <- schedtrim[(schedtrim$Date %in%datestouse), 4]
locationsched <- c("MatchLocationSingle", "NonMatchLocationSingle", "MatchLocationDouble", "MatchLocationDoubleEasy", "MatchLocationDoubleMust", "NonMatchLocationDouble", "NonMatchLocationDoubleEasy", "NonMatchLocationDoubleMust")
pretrainingsched <- c("PunishIncorrect", "PunishInc")
```

### Loop through files in raw directory and extract information on correct, incorrect, and correction trials.

```{r}
##Create empty data frame with named columns for data to be extracted
abet1keep <- data.frame(filesource = as.character(), totalcorrect = integer(), incorrect = integer(), corrtrials = integer(), correctcorr = integer(), centerincorrect = integer())
##Start schedule index at 1
schedindex <- 1

##Loop through filestouse one session at a time
    for(fil in filestouse) {
        ##Start with empty data frame formatted for 4 mice
        abet1trim <- data.frame(filesource = rep(as.character("TBD"),4), totalcorrect = rep(0,4), incorrect = rep(0,4), corrtrials = rep(0,4), correctcorr = rep(0,4), centerincorrect = rep(0,4)) 
        ##Read in abet generated csv file
        ##If session info is present, can change to skip = 14 to get everything to run
        abetdata1 <- read.csv(file = paste("../raw/",fil, sep=""), sep = ',', skip = 1, header = TRUE)
        ##Populate empty data frame with abet generated data
        abet1trim$filesource = as.character(fil)
        ##Pull different variables depending on schedule
        colindex <- 6
        for (rowindex in 1:4) {
            if (schedulestouse[schedindex] %in% pretrainingsched) {
            ##correct
            abet1trim$totalcorrect[rowindex]  <- as.vector(abetdata1[abetdata1$Item_Name == 
                    "Reward Collected Start ITI",colindex], mode = "numeric")
            ##incorrect
            abet1trim$incorrect[rowindex]  <-as.vector(abetdata1[abetdata1$Item_Name == 
                   "Incorrect",colindex], mode = "numeric")
            ##corrtrials obviously doesn't exist if schedule doesn't include corrections, set to 0
            abet1trim$corrtrials[rowindex]  <-as.vector(0, mode = "numeric")
            ##correctcorr also doesn't exist if schedule doesn't include corrections, set to 0
            abet1trim$correctcorr[rowindex]  <-as.vector(0, mode = "numeric")
            ##Increment counters to move to next session
            colindex <- colindex + 1
            schedindex <- schedindex + 1
            
            } else if (schedulestouse[schedindex] %in% locationsched) {
            ##correct
            abet1trim$totalcorrect[rowindex]  <- as.vector(abetdata1[abetdata1$Item_Name == 
                    "Reward Collected Start ITI",colindex], mode = "numeric")
            ##incorrect
            abet1trim$incorrect[rowindex]  <-as.vector(abetdata1[abetdata1$Item_Name == 
                   "Incorrect",colindex], mode = "numeric")
            ##corrtrials 
            if (sum(abetdata1$Item_Name == "Total_No_Correction_Trials") > 0) {
                abet1trim$corrtrials[rowindex]  <-as.vector(abetdata1[abetdata1$Item_Name == 
                   "Total_No_Correction_Trials",colindex], mode = "numeric")
            } else {
                abet1trim$corrtrials[rowindex]  <-as.vector(0, mode = "numeric")
            }            
            ##correctcorr
            if(sum(abetdata1$Item_Name == "Correction_Trial_Correct_Counter") > 0) {
                abet1trim$correctcorr[rowindex]  <-as.vector(abetdata1[abetdata1$Item_Name == 
                 "Correction_Trial_Correct_Counter",colindex], mode = "numeric")
                } else {
                   abet1trim$correctcorr[rowindex]  <-as.vector(0, mode = "numeric") 
                }
                
            ##Increment counters to move to next session
            colindex <- colindex + 1
            schedindex <- schedindex + 1            
            } else {
            ## If not one of the above schedules, set EVERYTHING except total correct trials to zero.  This is either Initial Touch, Must Touch, or some improperly named schedule that you should notice and fix.
            ##correct
            abet1trim$totalcorrect[rowindex]  <- as.vector(abetdata1[abetdata1$Item_Name == "_Trial_Counter",colindex], mode = "numeric")
            ##incorrect
            abet1trim$incorrect[rowindex]  <-as.vector(0, mode = "numeric")
            ##corrtrials
            abet1trim$corrtrials[rowindex]  <-as.vector(0, mode = "numeric")
            ##correctcorr
            abet1trim$correctcorr[rowindex]  <-as.vector(0, mode = "numeric")
            ##Increment counters to move to next session
            colindex <- colindex + 1
            schedindex <- schedindex + 1
            }
            }
            abet1keep <- rbind(abet1keep,abet1trim)
            
        }
    
    
##Bind schedule info to abet output
setalltrim <- cbind(schedtrim,abet1keep)
##Print bound info for first day
print(head(setalltrim,12))
#Print bound info for last day
print(tail(setalltrim,12))
```

### Calculate summary measures including percent correct for both correction and non-correction trials
```{r}
###Warning: It is possible to initiate a correction trial but not finish it, leading the total number of correction vs. non-correction trials to be off by 1.  This is unlikely to present a problem over large numbers of trials, but may impact sessions where mice did very few trials.  Any session where total correct responses are <5 should be viewed with great suspicion.  

##Set NA values to zero before you calculate summary measures
setalltrim[is.na(setalltrim)] <- 0
##totaltrials
setalltrim$totaltrials <- (setalltrim$totalcorrect + setalltrim$incorrect)
##reassign totaltrials to totalcorrect for DoubleMust schedules
setalltrim[setalltrim$Schedule %in% c("MatchLocationDoubleMust","NonMatchLocationDoubleMust"),11] <- setalltrim[setalltrim$Schedule %in% c("MatchLocationDoubleMust","NonMatchLocationDoubleMust"),6]
##noncorrtrials
setalltrim$noncorrtrials <- (setalltrim$totaltrials - setalltrim$corrtrials)
##reassign noncorrtrials to totalcorrect for DoubleMust schedules
setalltrim[setalltrim$Schedule %in% c("MatchLocationDoubleMust","NonMatchLocationDoubleMust"),12] <- setalltrim[setalltrim$Schedule %in% c("MatchLocationDoubleMust","NonMatchLocationDoubleMust"),6]
##incorrectcorr
setalltrim$incorrectcorr <- (setalltrim$corrtrials - setalltrim$correctcorr)
##firstcorr
setalltrim$firstcorr <- (setalltrim$totalcorrect - setalltrim$correctcorr)
##firstincorr
setalltrim$firstincorr <- (setalltrim$incorrect - setalltrim$incorrectcorr)
##corrperincorrect
setalltrim$corrperincorrect <- (setalltrim$corrtrials)/(setalltrim$firstincorr)
##percentall
setalltrim$percentall <- 100*(setalltrim$totalcorrect)/(setalltrim$totaltrials)
##percentfirst
setalltrim$percentfirst <- 100*(setalltrim$firstcorr)/(setalltrim$noncorrtrials)
##percentcorr
setalltrim$percentcorr <- 100*(setalltrim$correctcorr)/(setalltrim$corrtrials)

##Print summary measures for first day
print(head(setalltrim,12))
##Print summary measures for last day
print(tail(setalltrim,12))
```

### Export summary measure output to new csv file

Cleaned data with summary measures will be output to the output folder of the current working directory.  If this directory doesn't exist, you'll need to create it first.  Files created on the same DATE will be overwritten.

```{r}
write.csv(setalltrim, file = paste("../output/",Sys.Date(),".output.csv", sep=""), row.names = FALSE)
```

###Output graph of percent correct over all trials since beginning of training

```{r}
library(ggplot2)
library(RColorBrewer)
g<- ggplot(setalltrim, aes(x= Date, y = percentall, group = Mouse))
p<- g + geom_line(aes(color = Mouse), size = 1.5) + scale_colour_brewer(type = "qual", palette = "Paired") + geom_hline(aes(yintercept = 85), size = 1.5)
print(p)
```

###output graph of percent correct for just the most recent 3 days
```{r fig.width = 10, fig.height = 8}
last3days <- setalltrim$Date %in% tail(setalltrim$Date,36)
setall3 <- setalltrim[last3days,]
g<- ggplot(setall3, aes(x= Date, y = percentall, group = Mouse))
p<- g + geom_point(aes(color = Schedule), size = 4) + geom_line(aes(color = Schedule), size = 1.5) + scale_colour_brewer(type = "qual", palette = "OrRd") + geom_hline(aes(yintercept = 85), size = 1.5) + facet_wrap(~ Mouse)
print(p)
```

###output individual graphs of number of trials completed by mouse, colored by schedule
```{r fig.width = 10, fig.height = 8}
setalltrim$Schedule <- factor(setalltrim$Schedule, levels = c("Initial Touch", "Must Touch", "PunishInc",  "MatchLocationSingle", "NonMatchLocationSingle", "MatchLocationDouble", "MatchLocationDoubleEasy","MatchLocationDoubleMust", "NonMatchLocationDouble", "NonMatchLocationDoubleEasy", "NonMatchLocationDoubleMust"))
g <- ggplot(setalltrim, aes(x = Date, y = totaltrials)) + geom_point(aes(color = Schedule), size = 3) + scale_colour_brewer(type = "qual", palette = "Paired") + ggtitle("Total Trials") + coord_cartesian(ylim = c(0,105)) + facet_grid(Mouse ~ .) + facet_wrap(~ Mouse, ncol = 3)
print(g)
```

###output graph of percent correct for each mouse individually, colored by schedule
```{r fig.width = 10, fig.height = 8}
g <- ggplot(setalltrim, aes(x = Date, y = percentall)) + geom_point(aes(color = Schedule), size = 3) + scale_colour_brewer(type = "qual", palette = "Paired") + ggtitle("% Correct Over All Trials") + coord_cartesian(ylim = c(0,105)) + facet_grid(Mouse ~ .) + facet_wrap(~ Mouse, ncol = 3)
print(g)
```

###output correct FIRST TIME RESPONSES for each mouse, colored by schedule
```{r fig.width = 10, fig.height = 8}
g <- ggplot(setalltrim, aes(x = Date, y = percentfirst)) + geom_point(aes(color = Schedule), size = 3) + scale_colour_brewer(type = "qual", palette = "Paired") + ggtitle("% Correct Non-Correction Trials") + coord_cartesian(ylim = c(0,105)) + facet_grid(Mouse ~ .) + facet_wrap(~ Mouse, ncol = 3)
print(g)
```

###output total incorrect responses for each mouse, colored by schedule

Note that for DoubleMust schedules, the incorrect responses reflect only incorrect responses to the foil and not to the blank square.  On DoubleMust schedules, incorrect responses also do not result in the end of the trial.

```{r fig.width = 10, fig.height = 8}
g <- ggplot(setalltrim, aes(x = Date, y = incorrect)) + geom_point(aes(color = Schedule), size = 3) + scale_colour_brewer(type = "qual", palette = "Paired") + ggtitle("Total Incorrect Responses") + coord_cartesian(ylim = c(0,105)) + facet_grid(Mouse ~ .) + facet_wrap(~ Mouse, ncol = 3)
print(g)
```

###output total correct responses for each mouse, colored by schedule

```{r fig.width = 10, fig.height = 8}
g <- ggplot(setalltrim, aes(x = Date, y = totalcorrect)) + geom_point(aes(color = Schedule), size = 3) + scale_colour_brewer(type = "qual", palette = "Paired") + ggtitle("Total Correct Responses") + coord_cartesian(ylim = c(0,105)) + facet_grid(Mouse ~ .) + facet_wrap(~ Mouse, ncol = 3)
print(g)
```

###Display mean performance across blocks of 5 days
```{r}
setalltrim$Block <- as.factor(cut(setalltrim$DaysOnSchedule, c(1,6,11,16,21,26,31), include.lowest = TRUE, right = FALSE, labels = FALSE))
library(plyr)
blockedmeans<- ddply(setalltrim, .(Mouse, Schedule, Block), summarize, blockedmean = mean(percentall))
blockedmeansshort <- blockedmeans[(blockedmeans$Schedule != "Initial Touch" & blockedmeans$Schedule != "Must Touch"),]
g <- ggplot(blockedmeansshort, aes(x = Block, y = blockedmean, group = Mouse)) + geom_point(aes(color = Mouse), size = 4) +geom_line(aes(color = Mouse), size = 1.5) + geom_hline(aes(yintercept = 85)) + geom_hline(aes(yintercept = 50)) + scale_colour_brewer(type = "qual", palette = "Paired") + facet_grid(Schedule ~.) + facet_wrap( ~ Schedule, ncol = 3)
print(g)
```

